Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-07-03 manuscript/bef_stability/manuscript.Rmd
Finish review results, start discussion

---
title: Horizontal and vertical diversity jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, (Willem Bonnaffe ici?), Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

The effect of horizontal  and vertical diversity on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effect species richness and  network structure on total biomass, productivity
and stability. However, they have mainly been studied separately.  We used a
dataset of standardized fish sampling in French stream running from 1995 to
2018. We inferred trophic interactions thanks to recorded body size and diet
literature. We then investigate the direct and indirect effects of stream
size and temperature, horizontal and vertical diversity on total biomass and
its temporal stability. We found that horizontal and vertical diversity had
a positive relationship with total biomass while they had respectively a
negative and positive effect on temporal stability. We also found that
horizontal and vertical diversity are not independent and that environment
had an tremendous effect on total biomass and its temporal stability notably
through its effect on horizontal and vertical diversity. We conclude that our
study highlights that the effects horizontal and vertical diversity should not
treated separately and that it opens exciting perspectives on how environment,
horizontal and vertical can jointly determine total biomass in natural
communities.

(194/150 words)

@@

Current biodiversity losses and global changes affect ecosystems in complex and
unpredictable ways [@bellard_impacts_2012]. Since the seminal of paper of
@may_will_1972, the relationships between biodiversity and ecosystem properties
have received a lot of attention [@hector_plant_1999; @loreau_biodiversity_2001;
@hooper_effects_2005; @dunne_network_2006; @duffy_why_2009;
@isbell_forest_quantifying_2018]. Biodiversity largely determines both the
amount and the distribution of matter and energy fluxes within ecosystems, and
has hence a tremendous effect on ecosystem properties [@duffy_why_2009], such as
its total biomass [@hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016; @pennekamp_biodiversity_2018], biomass production
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016]
and the biomass temporal stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006;
@olivier_independent_nodate].

#### Horizontal and vertical diversity are two ways to look at biodiversity

Biodiversity can be quantified in numerous ways: either as the number of species
(i.e.  species richness) and in number of functional groups, which are usually
defined based on the trophic ecology of species, namely the way they acquire
food [@allan_stream_2007]. Those metrics can be organized into (i) "horizontal
diversity", which captures the overall diversity of components (i.e. species)
present in a community, and (ii) "vertical diversity" which captures diversity
of components across functional groups, such as the amount trophic categories,
often termed guilds or levels [@allan_stream_2007; @loreau_populations_2010;
@zhao_horizontal_2019; @wang_biodiversity_2016]. Vertical diversity has been
primarily quantified as food chain length [@loreau_populations_2010;
@zhao_horizontal_2019; @wang_biodiversity_2016] but it can also be characterized
by the food web structure [@wang_biodiversity_2016] such as connectance and
average trophic level. Horizontal and vertical diversity are, however, not
independent from each other. Connectance is known to covary negatively with
species richness [@winemiller_must_1989; @dunne_network_2006;
@thebault_stability_2010]. Species diversity also result in modification of
biomass distribution across the food web [@thebault_relationship_2006;
@duffy_ecosystem_2005; @duffy_functional_2007; @schneider_animal_2016].
Hence, it is crucial to study if horizontal and vertical diversity have synergic
or antagonistic effects and how their interplay modulates ecosystem properties.

#### horizontal and vertical diversity increase total biomass by common and different mechanisms

Horizontal diversity has been shown to promote higher biomass production in a
variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase primary productivity [@cardinale_species_2002,
@tilman_productivity_1996; @grace_integrative_2016]. Two mechanisms have been
proposed to rise this effect [@loreau_partitioning_2001]. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
Positive select effect arise when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
As for total biomass, empirical studies have generally demonstrated positive
relationships with species richness in plant communities [@hector_plant_1999;
@tilman_biodiversity_2014; but see @grace_integrative_2016] and mesocosms
[@pennekamp_biodiversity_2018]. When adding the vertical diversity, e.g.
consumers species, @gamfeldt_species_2005 and @thebault_relationship_2006 have found that total
biomass increased with increasing species diversity at two trophic levels.
But this positive relation seems to strongly depend on the niche differentiation of
the consumer species, i.e. their resource complementarity,
@thebault_relationship_2006. Hence, the total biomass-species richness
relationship seems to be supported by the same mechanisms within consumers and
within plants. @schneider_animal_2016 also showed in a allometric food web model
that species richness should increase total animal biomass.
Vertical diversity also affect total biomass. Theoretical studies showed that food
web properties, such as interaction distribution and strength, affect the
vertical distribution of biomass [@thebault_relationship_2006;
@schneider_animal_2016; @barbier_pyramids_2019], i.e. average trophic level.
More precisely, an increase of animal diversity results in
higher animal biomass as more biomass is transfered from plant to animals
[@schneider_animal_2016; @thebault_food-web_2003]. However, the role of vertical
diversity on the total biomass remains unclear in
natural communities.

#### Horizontal and vertical diversity versus temporal stability

Horizontal and vertical diversity also affects the stability of ecosystem
properties. Experimental studies showed consistently that species richness
increases the temporal stability of biomass in grasslands
[@tilman_biodiversity_2006], and mesocosms [@pennekamp_biodiversity_2018]. The
same relation was found in natural settings for temporal biomass stability in
stream fish communities [@franssen_annual_2011] and temporal abundance stability
of birds and butterflies [@olivier_independent_nodate, but not bats].  Two
mechanisms may govern the temporal stability of communities, (1) species
synchrony and (2) variability [@loreau_species_2008;
@thibaut_understanding_2013]. Species synchrony measures the temporal
covariation of species where low synchrony indicates that species temporal
variation compensate each other at the community level, thereby increasing
temporal stability. Species richness was found to decrease
synchrony between species [@loreau_species_2008; @thibaut_understanding_2013;
@bluthgen_land_2016; @olivier_independent_nodate; but @declerck_species_2006].
The species variability is measured as a average coefficient
of variation of abundance/biomass, weighted by species abundance/biomass. Higher species
temporal variability should thus result in lower temporal stability. Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. This conflicting patterns are dependent on
various mechanisms [see @thibaut_understanding_2013]. Species synchrony has been found to be
the main driver of temporal stability [@tilman_biodiversity_2006;
@olivier_independent_nodate] and responsible for the positive relationship
between horizontal diversity and temporal stability of biomass and abundance.

Vertical diversity, captured by network structure, has been at the heart of the
diversity-stability debate and largely explored by theoretical ecology
[@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. As species
interactions introduce inter-dependencies in species dynamics, they have a
substantial impact on stability. In particular, the strength of predation and
degree of omnivory in food webs have been shown to favor stability
[@neutel_reconciling_2007]. Furthermore, competition strength and niche breath of
consumers, can lead to negative or positive temporal stability-species richness
relationships [@thebault_trophic_2005] depending on the food web configuration
[@neutel_reconciling_2007].  However, it is important to note that theoretical ecology
generally defines stability based on community equilibrium and its stability to
small independent perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the amount
of secondary extinctions following species removal [@duffy_why_2009]. In
practice, these stability metrics hardly relate to empirical temporal stability
[@arnoldi_resilience_2016].

@@

Overall, the evidence suggests that both a higher horizontal and vertical
diversity should lead to a greater total biomass and stability in communities.
The effects of horizontal and vertical diversity are potentially conflicting and
interacting, as they rely on inter-dependent mechanisms, in such a way that
they need to be addressed jointly. Yet, there is no direct empirical evidence to
date for an interaction between effects of horizontal and vertical diversity on
total biomass and stability in natural communities.

@@

The environment affects organism performance directly, and thereby, species
abundances and total biomass [@grace_integrative_2016] and
temporal stability [@de_boeck_patterns_2017; @hansen_climate_2013;
@bluthgen_land_2016; @oliver_heterogeneous_2010; @friedpetersen_drivers_2020].
For instance, the hydrology of streams determines the composition of fish
communities, and is also related to fish population temporal stability
[@franssen_annual_2011]. Ecosystem size can also increase habitat heterogeneity
and so the number of different resource available, which in turn might increase
resource complementary between species and increase total biomass
[@post_ecosystem_2000; @doi_resource_2009]. Temperature has potentially
conflicting effects on horizontal and vertical diversity, as higher temperature
favour primary production and fluxes of biomass [@dossena_warming_2012;
@ogorman_chapter_2012], thus potentially supporting more species and trophic
levels [@arim_relationship_2007; @stegen_evolving_2012], but selects against
larger body sized individuals [@dossena_warming_2012; @edeline_ecological_2013;
@hattab_forecasting_2016], and thereby should reduce trophic diversity
[@hattab_forecasting_2016]. It is thus unclear whether these temperature effects
translate indirectly into changes in total standing biomass and stability in
natural food webs. The relative strength of direct and indirect effects of
environment on total biomass and its temporal stability remains an open
question.

#### Research questions

Overall, the joint effects of horizontal and vertical diversity on total biomass
and its temporal stability remains poorly documented, especially in natural
communities. In this study, we identified (1) the effects of horizontal
diversity and (2) vertical diversity on total biomass and temporal stability
(Figure \@ref(fig:matmet) A).  Considering both direct & indirect effects, (3)
we quantify if the magnitude of environmental effects on total biomass and its
temporal stability and compare them to that of horizontal and vertical
diversities.  (4) We identify the pathway for the effects of environment,
horizontal and vertical diversities on total biomass and its temporal stability.
(4a) Whether environmental effects operate mainly (i) directly, indirectly (ii)
through horizontal diversity, (iii) through horizontal and vertical diversities
or (iv) indirectly through vertical diversity only? (4b) Whether effects of
horizontal diversity arise (i) directly, or (ii) indirectly through vertical
diversity? Finally, (4c) whether environmental variables, horizontal and
vertical diversities affect species synchrony and species variability ?

We used the ONEMA dataset consisting in standardized sampling of fish
communities in french streams [@edeline_ecological_2013]. 99 sites were sampled for more
than 10 years spanning on the 1995-2018 period. We inferred
1496 interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow and enrichment. Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

@@

Stream fish community constitutes a good model to study the joint effects of
horizontal & vertical diversity on total biomass and its temporal stability.
Despite that fish trophic interactions are difficult to assess in nature, a consequent
body of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Most of the biomass-diversity and stability-diversity
relationships have been assessed in terrestrial ecosystems while there is a
evident need to test such relationships in a variety of ecosystem types. Stream
ecosystems are characterized by the flowing water and depends for a large part
from the input of terrestrial ecosystems. Furthermore, the biotic activity in
stream ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties.

We show that TODO

@@

Fish communities were monitored across stream sections in France on the period 1995-2018
by the French Office French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing, resulting in over XX individual fish sampled [see
@edeline_ecological_2013]. Two different standardized protocols were used for little and
big streams (detailed in Supplementary methods). For streams shallower than (XX
m), the sampling takes place over the whole width of the streams and is done by
foot whereas in streams deeper than (XX m), the sampling was done over the bank
only and by foot or by boat. The sampling took place either in late spring and/or Autumn
and temporal series were checked to reduce sampling heterogeneity within
stations (detailed in Supplementary methods). Each station have been followed
ten years or more. The temporal stability of biomass, computed as inverse of the
CV, can cover very different dynamics such as decreasing, stable or increasing.
However, the concept of community variability referred to variability around an
equilibrium, which is rarely observed or difficult to characterize in natural
communities. We finally kept only the 99 stations (Figure \@ref(fig:matmet) B)
which had no trends in biomass (method detailed in Supplementary methods).
However, the results were robust to the relaxation of this hypothesis
(Supplementary results).

##### Total biomass and its temporal stability

To characterize communities variability, abundance or biomass are generally
used. When the body size and therefore body mass vary by several orders of
magnitude inside a community, one downside of using abundance is that we mix
individuals that have potentially really different contributions in terms of ecosystem
properties such as total biomass and biomass repartition. We estimated
the body mass following an allometric law ($B_i = 0.01 \times
l_i^{3.03}$, $B$ being the body mass in grams, $l$ the body size in millimeters
and $i$ the individual) [@lleonart_removing_2000; @froese_cube_2006].

The species and total biomass were then computed for each sampling event.
As the sampling effort, reported as the surface sampled during the sampling event,
varied across protocol type and sampling event, biomass was divided by the
surface sampled (in $m^{2}$). So that, community and species biomass were expressed
in $gm^{-2}$.

In the analysis, the total biomass of a given station was characterized as the
median of the total biomass over time. Temporal biomass stability was defined as
the inverse of biomass variability. Variability was computed as Coefficient of
Variation (CV) of the biomass over time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and
$\mu_i$ being respectively the standard deviation and the temporal average of
the biomass of the station $i$). Then, temporal stability was equal to $CV^{-1}
= \mu \sigma^{-1}$.

To characterize the mechanisms driving temporal stability of communities, we
decomposed the community CV in synchrony and weighted average species CV
components according to @thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the Coefficient of Variation of the community in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
population in the station $i$ pondered by the biomass of each population.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_T}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$) [@loreau_species_2008]. $i$ being
the station $i$ and $k$ the species $k$.

$$
\overline{CV_{sp, i}} = \sum_k{\frac{B_k {CV}_i}{\sum_k{B_k}}}
$$

$$
\phi_i = \frac{\sigma^2_{x_T}}{(\sum_k{\sqrt{\sigma^2_{x_k}}})^2}
$$

## Horizontal & vertical diversity

### Horizontal diversity

@@

### Vertical diversity

@@

Trophic interaction network was inferred from body size and ontogenic diet.
Body size is a reasonable approximation widely used to infer trophic
interactions [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan. For example, alvins feed on plankton while adults may feed on
algae or fishes. Accordingly,  each fish species was divided in nine body size
classes. Each body size class constituting a trophic species. A high congruence
has been found at trophic species scale between inferred interactions and
measured trophic interactions, but not at the species
scale [@jennings_weak_2001]. We then inferred trophic interactions between the trophic species (Figure
\@ref(fig:matmet) C).  We determined network structure in two steps: (1) build a
metaweb describing the trophic interactions between all trophic
species (Figure \@ref(fig:matmet) C), then (2) determine network of a given
community by subtracting the metaweb according to the trophic species present in
a given community (Figure \@ref(fig:matmet) D). We added to each network seven
resource nodes present in food diet database: detritus, biofilm, phytoplankton,
zooplankton, macrophages, phytobenthos and zoobenthos (grey nodes in Figure
\@ref(fig:matmet) C & D). The trophic network inference procedure is detailed in
supplementary informations.

#### Trophic network structure

For each trophic network, corresponding to a particular community at a given
time and location, we computed two of the most communly used network metrics to
characterize network structure, namely the connectance and average trophic
level. Connectance describes the number of interactions compared to the number
of possible links, computed as ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). The average trophic level was
computed as the mean trophic level of the nodes that the predator nodes consumes
to which is added one, obtained recursively by setting the trophic level of
basal nodes to 1, and was further weighted by the biomass of each trophic
species (excluding resource nodes for which we had no biomass information)
($\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, $T_i$ and $B_i$
being resp. the trophic level and the biomass of the node $i$). In resume, we
infered trophic networks at the scale of trophic species (i.e. nodes being body
size classes) to take in account ontogenic diet shifts and that fish-fish
trophic interaction depends on body size ratio between prey and predator. On the
other side, the community metrics were defined at the species level, (i.e.
species richness, average species CV and synchrony).

@@

Habitat description, obtained from fish sampling data, describing the ecosystem
size namely the stream width, depth and position (altitude and
slope). Physico-chemical variables were obtained from
naiades web portal [http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/).
We chose flow and temperature to describe streams as flow
impose swimming constraints which filters fish species that can establish
[@rolls_scaling_2018; @tonkin_flow_2018; @gutierrez-canovas_contrasting_2013].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We took 
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF].
BOD is also inversely linked to oxygen concentration in water and so, BOD is
also linked to physiological stress of fishes. As physico-chemical variables
were not recorded at the same place than fish communities, they were
interpolated with spatial stream networks [@isaak_applications_2014;
@hoef_ssn:_2014]. The interpolation procedure is detailed in supplementary
methods.

>>>>

---
title: Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, (Willem Bonnaffe ici?), Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

The effect of species richness and network structure on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effect species richness and  network structure on total biomass, productivity
and stability. However, empirical data are lacking on the link between network
structure and ecosystem properties. We used a dataset of standardized fish
sampling in French stream running from 1995 to
2018. We inferred trophic interactions thanks to recorded body size and diet
literature. We then investigate the direct and indirect effects of stream
size and temperature, species richness and network structure on total biomass and
its temporal stability. We found that species richness and network structure had
a positive relationship with total biomass while they had respectively a
negative and positive effect on temporal stability. We also found that
species richness and network structure are not independent and that environment
had an tremendous effect on total biomass and its temporal stability notably
through its effect on species richness and network structure. We conclude that our
study highlights that the effects of species richness and network structure should be 
assessed conjointly and that it opens exciting perspectives on how environment,
species richness and network structure can jointly determine total biomass in natural
communities.

(210/150 words)

@@

Current biodiversity losses and global changes has the potential to affect
ecosystems in complex ways [@bellard_impacts_2012]. Biodiversity has a
tremendous effect on ecosystem properties because it largely determines both the
amount and the distribution of matter and energy fluxes
within ecosystems [@duffy_why_2009]. Biodiversity thus affects biomass
production [@tilman_productivity_1996; @cardinale_species_2002;
@grace_integrative_2016], biomass repartition [@thebault_relationship_2006] and
biomass stability [@thebault_trophic_2005; @thebault_relationship_2006;
@tilman_biodiversity_2006]. The relationship between biodiversity and
ecosystem properties is a central topic in ecology [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018] and its documentation is thus
crucial to better anticipate the impacts of global changes.

#### Species richness increases total biomass

Species richness promotes higher biomass production in a variety of ecosystems
and organisms. Indeed species richness generally increases primary productivity
and total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Theoretical models also
predict this pattern in plant communities [@tilman_plant_1997;
@loreau_partitioning_2001] and food webs [@schneider_animal_2016]. Resource use
complementarity and selection effects have been proposed to elucidate the
positive effect of species richness on productivity. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
A positive selection effect arises when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
In contrast, both positive [@wu_relationship_2014; @li_relationship_2018],
negative or no effects [@szwagrzyk_above-ground_2007; @jacob_productivity_2010;
@seidel_relationship_2013] of species richness on total biomass have been found
in natural forest. Some authors suggest that the lack of resource
complementarity or environmental harshness might be responsible for an absence or a
negative effect of species richness [@loreau_partitioning_2001; @he_global_2013;
@wu_relationship_2014]. Overall, under what conditions and through what
mechanisms we should observe positive relationships between biomass and species
richness in natural settings is still an open question.

Many studies showed that species richness also increases temporal stability,
i.e. variation through time, often quantified using CV. Such positive
relationships between species richness and temporal stability have been reported
in experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018]. Theoretical works
have clarified which mechanisms generate this positive relationship between
species richness and temporal stability [@loreau_species_2008;
@thibaut_understanding_2013]. Two mechanisms may govern the temporal stability
of communities, (1) species synchrony and (2) species variability. Species
synchrony quantify to what extent the fluctuations of abundances or biomasses
through time are correlated among species. A low synchrony in species
fluctuations, i.e.  asynchrony, is synonymous of compensatory dynamic, where the
decrease in abundance of a species is compensated by the increase of an other.
Asynchrony is generally the main mechanism whereby species richness has a
stabilizing effect on temporal stability [@tilman_biodiversity_2006;
@olivier_independent_nodate].  ^[Colin asked for a word about $cv_{sp}$ and
stability but my opinion is that it is not necessary and that the depth of mecanisms are
currently fairly explained for the introduction]. Such positive relationships
between speciesrichness and temporal are also observed in natural settings, such
as stream fish communities [@franssen_annual_2011], bird and butterfly
communities [@olivier_independent_nodate].

#### Multitrophic communities

One has to notice that most of the empirical and experimental evidence of the
positive effect of richness on biomass production and stablity through time
comes from studies focusing on one trophic level, generally primary producers or
primary consumers. Natural ecosystems generally include
more than one or two trophic levels and it is not certain that what holds
on one trophic level hold also in multitrophic communities.
Theoretical studies showed that food web properties, such as interaction
distribution and strength, affect the distribution of biomass among adjacent
trophic levels, [@thebault_relationship_2006; @schneider_animal_2016;
@barbier_pyramids_2019]. A
large body of theoretical studies links the propagation of perturbations to the
foodweb structure [@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011],
or trophic chain [@shanafelt_david_w._stability_nodate; @wang_biodiversity_2016;
@barbier_pyramids_2019]. To date, we have few evidence on the relationship
between network structure and total biomass and its temporal stability in
empirical multitrophic communities.

@@

Overall, the evidence suggests that both species richness and network structure
should modulate total biomass and stability of communities. The effects of
species richness and network structure are potentially conflicting or
synergistic in such a way that they need to be addressed jointly. Yet, there is
no direct empirical evidence to date for the combined effects of species
richness and network structure on total biomass and stability in natural
communities.

@@

When assessing the relationship between community structure and ecosystem
properties in natural communities, it is crucial 
to consider environmental factors [@loreau_biodiversity_2001;
@he_global_2013; @grace_integrative_2016]. Environment affects organism
performance directly, and thereby, species abundances, total biomass
[@grace_integrative_2016] and temporal stability [@de_boeck_patterns_2017;
@hansen_climate_2013; @bluthgen_land_2016; @oliver_heterogeneous_2010; @franssen_annual_2011;
@friedpetersen_drivers_2020]. For example, greater ecosystem size might increase
the number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009]. @li_relationship_2018 have notably shown that climate
moisture has a positive direct effect on total biomass, but also indirect
through a positive effect on species richness. Taking in consideration
environment also allow for an overall comparison of the strength of the
the effects of environment and community structure on total biomass and its
stability.

@@

Stream fish community constitutes a good model to study the effects of
network structure on total biomass and its temporal stability. Despite that
fish trophic interactions are difficult to measure in nature, a consequent body
of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Especially in stream communities, including fishes,
@brose_predator_2019 found that the prediction of predator prey trophic
interactions based on body size ratio was far better in stream ecosystems than
in other aquatic and te ecosystems. Furthermore, the biotic activity in stream
ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties. Lastly, most of the diversity-biomass and diversity-stability
relationships, have been assessed in terrestrial ecosystems
while there is a evident need to test such relationships in a variety of
ecosystem types.

We used the ONEMA dataset consisting in standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. 99 sites were sampled
for more than 10 years spanning on the 1995-2018 period. We inferred
1496 interaction networks, based on body size and food diet.  The network
structure was summarized by connectance and mean trophic level. Environment
was characterized by the stream size, altitude, slope, and the temperature, flow
and enrichment. Finally total biomass was measured as median total biomass over
the sampling period while its temporal stability was quantified as the inverse
of the CV.

We found a positive relationship between total biomass and species richness and
network structure  while total biomass and species richness had respectively a
negative and positive effect on temporal stability. We also showed a link
between species richness and network structure. We found that the environment had
a strong effect on total biomass and its temporal stability in particular by 
its effect on species richness and network structure. We then discuss the
possible mechanisms at play to explain those results and their implications on
our knowledge of the effect of network structure on ecosystem properties.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems
(ONEMA) using electric fishing. Two different standardized protocols were used
for little and big streams (detailed in Supplementary methods). For streams
shallower than (XX m), the sampling takes place over the whole width of the
streams and is done by foot whereas in streams deeper than (XX m), the sampling
was done over the bank by foot or by boat. The sampling season spans from 
late spring to Autumn. During sampling, all the fishes caught were
identified and counted. For each species, either the body size of all fish was
measured, or a representative sample when the number of fish was high (See
supplementary methods for details). The fishes were released at the end
of the sampling. Additionally, the physical characteristics of the river was
recorded during sampling as well as the sampling effort, quantified as the
surface sampled.

#### Data selection

We kept only the most frequent sampling protocol in each station and kept
stations that had been sampled ten years or more. The median sampling month was
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$, i.e. there was low heterogeneity of the sampling month
within stations. As the concept of community variability referred to variability
around an equilibrium, which is rarely observed or difficult to characterize in
natural communities, only those stations with no temporal trends in total
biomass were selected (detailled in supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results). We
retained 99 stations distributed over seven hydrographic basins (Figure
\@ref(fig:matmet) B).

## Data preparation

#### Biomass

The body size of the unmeasured individuals was inferred under the assumption
that species body size followed a normal distribution whose the mean and the
standard deviation was quantified from the measured individuals of the sampling
(See supplementary methods). The body mass of each individual was then estimated
using an allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$ being the body mass
in grams, $l$ the body size in millimeters and $i$ the individual), where the
coefficients $0.01$ and $3.03$ comes from meta-analysis
[@lleonart_removing_2000; @froese_cube_2006]. The species biomass
and total biomass at each sampling was then calculated by summing individual
body masses. To account for the sampling effort, species biomass and total
biomass have been related to the surface sampled, i.e. $gm^{-2}$.
In the further analysis, the total biomass of a station has been defined as the
median of the total biomass over time.

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the biomass of
the station $i$). Then, temporal stability was equal to $CV^{-1} = \mu
\sigma^{-1}$.

In order to describe the mechanisms driving temporal stability of biomass, we
decomposed the $CV$ of total biomass ($CV_{com}$) in synchrony ($\phi$) and
weighted average species $CV$ (\overline{CV_{sp}}) components according to
@thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$) [@loreau_species_2008]. $i$ being
the station $i$ and $k$ the species $k$.

$$
\overline{CV_{sp, i}} = \sum_k{\frac{B_{k,i} {CV}_{k, i}}{\sum_k{B_{k,i}}}}
$$

$$
\phi_i = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sqrt{\sigma^2_{x_{k,i}}}})^2}
$$

@@

### Network structure

@@

Trophic interaction network was inferred from body size and ontogenic diet shift.
Body size is a reasonable approximation widely used to infer trophic
interactions [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan. For example, alvins feed on plankton while adults may feed on
algae or fishes. Accordingly, each fish species was divided in nine body size
classes, each of which is trophic species (Figure SXX). A high congruence
between inferred interactions and measured trophic interactions has been found
at trophic species scale but not at the species scale [@jennings_weak_2001].
Additionally to the fish trophic species, we added seven resource nodes present
in ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Figure \@ref(fig:matmet)
C & D). The ontogenic diet database was based on informations from fishbase and
literature. Each fish species had two or three life stages which are delimited by
body size (Table SXX, Figure SXX).

The trophic interactions of fish trophic species was determining with the median
of its body size range, hereafter after midpoint. The midpoint of trophic
species fall into one of the species life stage and a trophic link was set
between the trophic species and the resources present in the life stage. If the
trophic species is in a piscivorous life stage, the fish-fish trophic links was
defined in two steps and based on the body size ratio between predators and
prey. First of all, we defined the range predation window from 3% to 45% of the
midpoint of the trophic species. Then, a trophic link was set between the
piscivorous trophic species and every trophic species whose midpoint was
included in the predation range (Figure SXX). Lastly, the trophic links between resources was
set according to the food diet data.

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Figure
\@ref(fig:matmet) C), then (2) determine network of a given community by
subtracting the metaweb according to the trophic species present in a given
site and sampling event (Figure \@ref(fig:matmet) D).

For each trophic network, corresponding to a particular site and sampling, we
computed two of the most commonly used network metrics to characterize the food
web, namely the connectance and mean trophic level. Connectance describes 
the level of generalism in the network and is calculated as the number of actual
interaction divided by the maximum number of interaction if all species interact
together ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). The trophic level of the
nodes was computed of the nodes that the predator nodes consumes
to which is added one, obtained recursively by setting the trophic level of
basal nodes to 1. The mean trophic level $\overline{T}$ was computed as the mean trophic level
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, with $T_i$ and $B_i$
being resp. the trophic level and the biomass of the node $i$.

@@

we characterised the sampled sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow as they affect the type of fish that can
establish [@franssen_annual_2011; @rolls_scaling_2018; @tonkin_flow_2018;
@gutierrez-canovas_contrasting_2013] as well as BOD (Biological Oxygen Demand)
as a proxy for eutrophisation. As data from the naide database are not collected
on same site as our fish sampling sites, we performed an interpolation with
spatial stream networks [@isaak_applications_2014; @hoef_ssn:_2014]. The
interpolation procedure is detailed in supplementary methods.

We had one environmental variable value for each site and sampling event, except
for water temperature for which the record began in 2006. Then, we computed the average over
time and the CV over time for water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA)
was performed on the 14 scaled variables. We kept in the analysis the two first
principal components based on eigen values distribution (Fig. SXX).
The two first principal components explained `r tot_var_expl_pca`%
of the variance. A varimax axis rotation was performed to maximize the representation
of the variables by the principal components (Figure SXX).
The two first principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature. Further description of the variables are detailed in
supplementary information.
